## **Q1**

*Lookup tables* are a common use case for named vectors. Let\'s say you\'re surveying the avian community of a wetland. In your notebook, you recorded species by their 4-letter code: GREG for Great Egret, MALL for Mallard, MAWR for Marsh Wren, and KILL for Killdeer. You need a table of counts, but you need the table to show the full common name, not just the 4-letter code you used for convenience. Modify the code below to accomplish that task. First, make `species_code` a named vector, then modify the call to `table()` to use `species_code` as a lookup table that converts 4-letter codes to common names.

```         
species_codes <- c( "GREG" = "Great Egret", "MALL" = "Mallard", "MAWR" = "Marsh Wren", "KILL" = "Killdeer")

sightings = c("GREG", "GREG", "MALL", "MAWR", "KILL", "GREG")

# Modify this line to use species_codes so your counts have common names
table(species_codes[sightings])
```

## **Q2**

These questions are about a built-in R class that is an extremely common source of confusion: `POSIXct`.

-   Explain why adding 1 to a `POSIXct` datetime object increments it by a second. What *type* is `POSIXct` (type, not class)? How does `POSIXct` use that type to represent datetimes?\
    \
    *POSIXct datetime objects are vectors "built on top of double vectors", where the value represents the \# of seconds since an origin datetime (e.g. 1970-01-01). Thus, by adding 1 to a datetime, you are adding a second since the units of the vector are seconds.*

-   What\'s the default **origin** of POSIXct objects? Write a line of code that converts the number 0 to January 1, 2000 at noon in UTC using `as.POSIXct()` (hint: look up the documentation for the `origin` parameter in the `as.POSIXct()` help page).

    *The default origin of POSIXct objects is 1970-01-01. The following code converts the origin to* *January 1, 2000.*

    ```{r}

    zero <- 0
    zero_datetime <- as.POSIXct(zero, 
                                origin = "2001-01-01 00:00", 
                                tz="UTC")
    zero_datetime
    ```

    -   Take away the POSIXct class from the object you just made. What kind of vector is it now? Why isn\'t the value 0 anymore?

        *The value is no longer 0 because we removed the reference "origin" value. This number represents the number of seconds from 1970-01-01.*

```{r}
zero_datetime <- as.vector(zero_datetime)
zero_datetime
```
